---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(ggplot2)

temp <- tempfile()
download.file("https://github.com/Moorecj87/RepData_PeerAssessment1/raw/master/activity.zip", temp)
dat <- read.csv(unzip(temp, "activity.csv"))

#convert interval to repeat each day

dat$interval <- rep(1:288, length.out = 17568)

```

## What is mean total number of steps taken per day?

```{r}

dat2 <- aggregate(steps ~ date, dat, FUN = sum)
hist(dat2$steps)

```

The mean is `r as.integer(mean(dat2$steps))` and the median is `r as.integer(median(dat2$steps))`.

## What is the average daily activity pattern?

```{r}

dat3 <- aggregate(steps ~ interval, dat, mean)

with(dat3, plot(interval, steps, type = "l"))

maxsteploc <- which.max(dat3$steps)
maxint <- dat3$interval[maxsteploc]

```

The interval with the maximum average number of steps is `r maxint`.

## Imputing missing values

```{r}

nasum <- sum(is.na(dat$steps))


```

The number of NA values is `r nasum`.

```{r}

# impute missing data using average per interval

# add average per interval to original dataset

dat4<-dat3
for(i in 1:60){dat4<-rbind(dat4, dat3)}
dat5<-cbind(dat, dat4$steps)

#replace NA with average per interval, remove unneccessary column

dat5$steps[is.na(dat5$steps)] <- dat5$`dat4$steps`[is.na(dat5$steps)]
dat5 <- dat5[,1:3]

dat6 <- aggregate(steps ~ date, dat5, FUN = sum)
hist(dat6$steps)

```

The means with NA vs without are `r as.integer(mean(dat2$steps))` and `r as.integer(mean(dat6$steps))` repectively.  The medians are `r as.integer(median(dat2$steps))` and `r as.integer(median(dat6$steps))` respectively.

## Are there differences in activity patterns between weekdays and weekends?

```{r}

dat7 <- mutate(dat5, newdate = weekdays(as.Date(dat5$date)))

weekdaylist <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
dat7$newdate <- factor((dat7$newdate %in% weekdaylist), levels = c(FALSE,TRUE),                              labels = c('weekend','weekday'))

dat7.weekday <- aggregate(steps ~ interval, dat7[dat7$newdate=="weekday",], mean)
dat7.weekend <- aggregate(steps ~ interval, dat7[dat7$newdate=="weekend",], mean)

par(mfcol=(c(1,2)))
plot(dat7.weekday, type = "l", main = "Weekdays")
plot(dat7.weekend, type = "l", main = "Weekends")

```